---
title: "Leaflet Maps"
author: 'Visualization Workflows'
date: 'ENVS543: Environmental Data Literacy'
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    hightlight: tango
    logo: "../../logo.png"
    css: "../../style.css"
---

```{r setup, include=FALSE}
source("../../../options.R")
```



## Leaflet

A leaflet map has several components, just like a `ggplot` object, which are layered ontop of eachother. 

 - Javascript library
 - Integrated into R 
 - Interactive mapping on HTML


```{r}
library(leaflet)
library(magrittr)
```


# Basic Mapping


## Basic Plotting

```{r}
leaflet() %>% addTiles()
```



## Alternative Tile Providers: ESRI


```{r}
leaflet() %>% 
  addProviderTiles(provider="Esri.WorldStreetMap") %>%
  addMarkers(lng=-77.454459, lat=37.544700)
```


## Tile Providers: Open Maps

```{r}
leaflet() %>% 
  addProviderTiles(provider="OpenTopoMap") %>%
  addMarkers(lng=c(-77.45,-81.63), lat=c(37.54,38.34) )
```

### Artistic Maps

```{r}
leaflet() %>% 
  addProviderTiles(provider="Stamen.Watercolor") %>%
  addMarkers(lng=c(-77.45,-81.63), lat=c(37.54,38.34) )
```

## Markers

```{r echo=FALSE}
df <- data.frame( University = c("Seatle Central Community College",
                             "Western Washington University",
                             "University of Missouri - Saint Louis",
                             "Iowa State University",
                             "Virginia Commonwealth University"),
                  State = c("Washington",
                            "Washington",
                            "Missouri", 
                            "Iowa", 
                            "Virginia"),
                  Reason = c("AA Degree",
                             "BS Degree in Botany","MS/PhD in Biology",
                             "PostDoctoral Researcher",
                             "Professor (e.g., Get a Job!)"),
                  Latitude =  c(  47.612,  48.734, 38.708, 42.025, 37.544),
                  Longitude = c(-122.323,-122.488,-90.307,-93.646,-77.453)
                  )
df %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```



## Dialog Boxes

```{r}
popup = paste( "<strong>", df$University,"</strong>",
                 "</br>Reason for Attending: <i>", df$Reason, "</i>", sep="")
popup
```


## Integrating into Markers

```{r}
df %>% leaflet() %>% addTiles() %>% addMarkers( popup = popup)
```



---

From [Iowa State University Envrionment Mesonet](https://mesonet.agron.iastate.edu/ogc/), we can grab WMS (Web Map Service) Tiles.  This is the most recent reflectance (when the slides were compiled).

```{r}
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 4) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "<br>Weather data © IEM Nexrad"
  )
```




# Abstract Geometries

## Mapping Polygons

```{r}
library(maps)
states <- map("state",fill=TRUE, plot=FALSE)
summary(states)
```



## Mapping Polygons


```{r}
leaflet( states ) %>% addTiles() %>%
  addPolygons( fillColor= rainbow(10, alpha=NULL), stroke=TRUE, weight=1.0 )
```


