---
title: "GGPlot Homework"
author: "Put Your Name Here"
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, echo=FALSE}

# This code is the 'setup' chunk.  An it is run each time you knit the document.  What the code
#  below does is check to see if you have the DT library installed on the machine you are working
#  on, and if you do not, it installs it for you.
if( !require(DT) ) {
  install.packages("DT")
}

# I am also going to load in the tidyverse constellation of packages that all have the 
#   ggplot and other libraries for us.  If we put it in the setup chunk, we don't get 
#   that huge dump of messages in the document AND we are assured that each time we open
#   and run the code in this notebook, all the correct libraries are loaded in for us.
library(tidyverse)
```


## The Data

This homework will focus on developing graphing skills using the `ggplot2` library.  For this, we are going to use an existing dataset named `CO2` consisting of 84 records of cold tolerance for the grass species *Echinochloa crus-galli*.  The data itself consists of the rate of carbon dioxide uptake for a dozen plants, divided amongst two treatments, the first is chill/no-chill & and the second one represents the origin of the plants (Quebec & Mississippi).  The raw data columns are:  
  - Plant An ordered `factor` indicating the sample plant.  
  - Type  An unordered factor indicating the origin of the plant.  
  - Treatment  An unordered factor representing the cold treatment.  
  - conc  The concentration of ambient carbon dioxide.  
  - uptake  The rate of updake of ambient carbon dioxide $\left[ \frac{\mu M}{m^2} * sec \right]$.  

Which in R, the raw data look like:

```{r}
# load in the DT (data.table) library (installed in the setup chunk) has the
#  datatable function that allows you to ahve an interactive widget in HTML form
#  that allows you to see and make queries of the data set.
library(DT)

datatable( CO2 )
```

and can be summarized as:

```{r}
summary(CO2)
```

## Summarizing Univariate Data

For the first set of plots, we will summarize a single set of data, `uptake`. In the following chunks, comments indicate what kind of graphics should be made using `ggplot` functions.  I load in the `tidyverse` constellation of libraries (of which `ggplot2` is one of) and we can go from there.



```{r}
# create a histogram of CO2 uptake

```





## Summarizing Bivariate Data

For this section, we are going to focus on two sets of data.




## Colors and Shapes










