---
title: "Review of Section 1"
author: "Visualizations"
date: "2/20/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review

This is an overview of the content that was covered thus far in Data Literacy.


## Base Skills


### Getting Data

You can make data directly, by entering it into R.

```{r}
x <- c( 1, 2, 6, 3, 2, 7, 4, 2, 6 )
y <- c( 9, 3, 6, 2, 6, 3, 2, 6, 3 )
```

The fundamental object we keep data in is the `data.frame`.

```{r}
df <- data.frame( X = x, 
                  Y = y )
df
```

There is data that is already include in `R` for examples.  The `iris` data set is an example.

```{r}
summary(iris)
head(iris)
```


We can also load data in from a local file.

```{r}
# I'm assuming that thre is is a folder called 'data' and within that folder there
#  is a file called 'araptus.csv'  It was included with a lecture.

araptus <- read.csv("data/araptus.csv")
summary( araptus )
```


## Basic Plotting

`R` comes with some basic plotting command that can make a wide array of graphics (though most seem to be from the 90's in terms of overall style).  Here are some examples.

```{r}
# histogram of suitability scores for observed araptus sampling locales
hist( araptus$Suitability, 
      xlab="Habitat Suitability", 
      ylab="Frequency", 
      breaks = 10,
      main = "Araptus Habitat Suitibility")
```

```{r}
# Scatter plot female and male beetle census numbers
plot(  araptus$Males ~ araptus$Females, 
       xlab ="Female Beetles",
       ylab = "Male Beetles" )
# add a x=y line to indicate where equal ratios of male and females would be
#   the a= is the intercept, the b= is the slope, col is the color and lty is 
#   the line type.
abline( a=0, b=1,
        col = "red", 
        lty = 2)
```

You can customize the *p*lot *ch*aracters using `pch`.  There are 25 recognized characters.

```{r}
plot( 1:25,
      1:25,
      pch = 1:25 )
```




```{r}
boxplot( iris$Sepal.Length ~ iris$Species, 
         notch = TRUE, 
         xlab = "Iris species",
         ylab = "Sepal Length" )
```

```{r}
# this is a barplot
mean_petal_length <- c(  1.46, 4.26, 5.52 )
barplot( mean_petal_length,
         names.arg = c( "Iris setosa",
                        "Iris versicolor",
                        "Iris virginica"),
         xlab = "Species",
         ylab = "Average Petal Length (cm)")
```

You can even make bad dad jokes with these plots...

```{r}
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```



## GGPlot

The `ggplot2` library allows you to overlay layers, geometries, transformations, and statistics onto more modern-looking graphical output.  At a base level, we need:

 1. To make a plot using `ggplot()`
 2. Define the data being used.
 3. Specify an `aes`ethic which lets the plot know which parts of the data you are going to use to plot or color components within the each layer.
 4. Add `geom`etric layers to the plots.
 5. Make labels, etc.

Using the `airquality` data.

```{r}
data(airquality)
summary( airquality )
```

we can look at univariate data distributions (e.g., a single variable)
```{r, warning=FALSE, message=FALSE}
library( ggplot2 )
ggplot( airquality, aes( Solar.R) ) +
  geom_histogram()
```

Or make plots of two variables such as scatter plots

```{r}
ggplot( airquality, aes( Temp, Wind ) ) +
  geom_point() +
  xlab("Temperature (F)") + 
  ylab("Windspeed (MPH)")
```

even ones with labels.

```{r}
kite.days <- data.frame( Wind = c(20.7,20.1),
                         Temp = c(72, 61),
                         Labels = c("Best Day", "Next Best Day")
)
kite.days
```

(and showing how to use two different data.frame objects to plot on the same plot)

```{r}
library( ggrepel )
ggplot( airquality, aes( Temp, Wind ) ) + 
  geom_point() + 
  geom_label_repel( aes(label = Labels), data = kite.days )
```

even with colors

```{r}
ggplot( airquality, aes(Temp, Wind, color=Month)) + 
  geom_point() +
  ylab("Windspeed") + 
  xlab("Temperate (F)") 
```


or boxplots where the x-axis data is 'categorical'

```{r}
ggplot( airquality, aes(Month, Wind) ) + 
  geom_boxplot( notch=TRUE)
```



## Rasters

We can treat a raster just like any other object and use it in displaying geospatial content.

```{r message=FALSE, warning=FALSE}
library(raster)
r <- raster("data/alt_22.tif")
r
```

To do a variety of things such as defining an extent within the raster

```{r}
e <- extent( -116, -108, 22, 30)
e
```

and cropping it (compare the extent of this one to the previous one).

```{r}
baja <- crop( r, e )
baja
```

and plot

```{r}
plot( baja, 
      xlab = "Longitude",
      ylab = "Latitude", 
      main = "Elevation for Baja California",
      col = terrain.colors( 50 ) )
```

We can even break individual rasters up and make them `data.frame` objects so that we can use `ggplot` and all of its glory to plot them.

```{r warning=FALSE, message=FALSE}
library( magrittr)

raster("data/alt_22.tif") %>%
  crop( extent( -116, -108, 22, 30) ) %>%
  rasterToPoints() %>%
  data.frame() -> baja_df

summary( baja_df )
```

and overlay points

```{r}
baja_df %>%
  ggplot() + 
  geom_raster( aes(x,y,fill=alt_22)  ) + 
  coord_equal() + 
  scale_fill_gradientn( name="Elevation (m)", 
                        colours = terrain.colors(100)) + 
  theme_bw() + 
  xlab("Longitude") + ylab("Latitude") +
  geom_point( aes(x=Longitude,y=Latitude, size=Suitability), 
              data = araptus, 
              col = "red",
              alpha = 0.4)
```

We can `extract` data from a raster for an analysis by defining the points.

```{r}
pts <- SpatialPoints( cbind( araptus$Longitude, araptus$Latitude ) )
site_elev <- extract( baja, pts)
site_elev
```

for subsequent analysis

```{r}
araptus$Elevation <- site_elev
ggplot( araptus, aes( Latitude, Elevation ) ) + 
  geom_line() + 
  theme_bw()
```


## Leaflet

We can use the leaflet library to create interactive maps.  Thus far, we've worked on making individual maps with points and markers.

```{r fig.cap="*Figure 1:* The spatial distribution of sample locations for the Sonoran Desert bark beetle (*Araptus attenuatus*).",  warning=FALSE, message=FALSE}
library(leaflet) 

leaflet(data = araptus) %>%
  addProviderTiles(provider="Stamen.TerrainBackground") %>%
  addCircleMarkers(weight=1, fill=TRUE, color="#fff") %>%
  addLabelOnlyMarkers(label=~as.character(Site), 
                      labelOptions = labelOptions(noHide=TRUE,
                                                  direction='center',
                                                  textOnly = TRUE))
```

We can even get more fancy and overlay things like charts

```{r}
library(leaflet)
library(leaflet.minicharts)

leaflet() %>%
  addProviderTiles( providers$Esri.WorldTopoMap) %>%
  addMinicharts( 
    araptus$Longitude, araptus$Latitude,
    type = "pie",
    chartdata = araptus[, c("Males","Females")],
    width = araptus$Suitability*30
    )
```

and layers provided by *W*eb *M*apping *S*ervices such as weather layers.

```{r}
leaflet() %>% 
  setView( lng=-86.0490, 35.1495, zoom = 5 ) %>% 
  addProviderTiles( providers$OpenTopoMap ) %>% 
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = c(WMSTileOptions(format = "image/png", transparent = TRUE),
                providerTileOptions( opacity = 0.45) ),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```
