---
title: "Bootcamp - R, RStudio, Interactions"
output: html_document
---

```{r setup, include=FALSE}
source("./startup.R")
```


> This is a short outline of the materials and R content that is presented as the 'bootcamp' portion of the course.


# Background 

This is the stuff that needs to be gone over before we can start being interactive.  

> ... Robert Gentleman and I became colleagues at The University of Auckland. We both had an interest in statistical computing and saw a common need for a better software environment in our Macintosh teaching laboratory. We saw no suitable commercial environment and we began to experiment to see what might be involved in developing one ourselves.

Which resulted in the following email being posted to the s-news listserve on 4 August 1993 by Ross Inhatka and Robert Gentleman in response to a question about getting S running on macintosh computers (they just came out and were quite popular at the time).


> Date: Wed, 4 Aug 93 14:01:31 NZS  
> From: Ross Ihaka   
> To: S-news@utstat.toronto.edu  
> Subject: Re: Is  S  available for a Macintosh personal computer?  
> 
> Joseph B Kruskal writes:
> 
> > If anyone knows of an S available for a Macintosh computer, I
> would be pleased to hear about it. 
> 
> About a year ago Robert Gentleman and I considered the problem of
> obtaining decent statistical software for our undergraduate
> Macintosh lab.  After considering the options, we decided that
> the most satisfactory alternative was to write our own.  We
> started by writing a small lisp interpreter.  Next we expanded
> its data structures with atomic vector types and altered its
> evaluation semantics to include lazy evaluation of closure
> arguments and argument binding by tag as well as order.  Finally
> we added some syntactic sugar to make it look somewhat like S.
> We call the result "R".
> 
> R is not ready for release at this point, but we are committed to
> having it ready by March next year (when we start teaching with
> it).
> 
> Because there is likely to be some interest in it we are going to
> put some (SPARC/SGI/Macintosh) binaries on Statlib later this
> week so people can give it a test drive.
> 
> 'll send out a short note when the binaries are available.
> 
> THIS IS NOT AN OFFICIAL RELEASE.  PLEASE DON'T BUG US ABOUT IT.
> WE ARE FULLY OCCUPIED WITH TEACHING AND ADMINISTRATIVE WORK AT
> PRESENT.
> 
> WHEN THE SOURCE CODE IS STABLE, WE WILL MAKE IT AVAILABLE THROUGH
> STATLIB.
> 
> Robert will be at the ASA meetings in S.F. next week ...
>
> ...
> 
> Ross Ihaka


## Chronology


```{r fig.cap="A short chronology of S and R.", fig.lp="fig"}
source("./1_Bootcamp_Chronology.R")
```


## Data & Narrative

What if we can mix the data

```{r}
araptus <- read_csv( "araptus.csv")
datatable( araptus )
```



## Data & Narrative 

The spatial distribution of $N =$ `r sum(araptus$Male) + sum(araptus$Female)` individuals were sampled from $K =$ `r nrow(araptus)` sampling sites covered the entire known extent for *Araptus attenuatus* (Figure 1).  On average the sex ratio (Females:Males) within each of the sampling locales varied considerably, ranging from a low of `r min( araptus$Females / araptus$Males )` to `r max( araptus$Females / araptus$Males )`.

```{r fig.cap="*Figure 1:* The spatial distribution of sample locations for the Sonoran Desert bark beetle (*Araptus attenuatus*).", eval=TRUE, fig.align="left"}
library(leaflet) 

leaflet(data = araptus) %>%
  addProviderTiles(provider="Stamen.TerrainBackground") %>%
  addCircleMarkers(weight=1, fill=TRUE, color="#fff") %>%
  addLabelOnlyMarkers(label=~as.character(Site), 
                      labelOptions = labelOptions(noHide=TRUE,
                                                  direction='center',
                                                  textOnly = TRUE))


```

Evidence of differential sex-biased dispersal is revealed by differential structure in biparental and uniparentally inherited genetic markers.  The expectation under equal dispersal betweeen sezes for maternally inherited mitochondrial markers ($F_{ST;mt}$) and biparent nuclear markers ($F_{ST; nuc}$) are (see also Hartl & Clark 1997; Wang 1997 for another formulation):

\[
F_{ST; mt}= \frac{8 F_{ST;nuc}}{1 + 7 F_{ST; nuc} }
\]



```{r echo=TRUE}
library(leaflet)
library(leaflet.minicharts)

leaflet() %>%
  addProviderTiles( providers$Esri.WorldTopoMap) %>%
  addMinicharts( 
    araptus$Longitude, araptus$Latitude,
    type = "pie",
    chartdata = araptus[, c("Males","Females")],
    width = araptus$Suitability*30
    )

```


## Results

```{r}
araptus %>%
  mutate( SexRatio = Females / Males ) -> araptus

model <- glm( SexRatio ~ log(Suitability), 
              data = araptus,
              family = Gamma() )
summary(model)
```

Female beetles seem to be selecting locales in more suitable habitats (Figure 2), which is consistent with a Gaussian log normal distribution (GLM; intercept = `r summary(model)$coefficients[1,1]`; coef =`r summary(model)$coefficients[2,1]`; P =`r summary(model)$coefficients[2,4]`).  This suggests that males who establish brood chambers at the margins of the species niche are actively selected against by maturing females.

```{r fig.cap="*Figure 2:* Variation in the observed sex ratio of *A. attenuatus* locales as a function of habitat suitability as determined by MAXENT Niche modeling of host plant distributions.", fig.align="left"}
araptus %>%
  ggplot( aes(Suitability, SexRatio) ) + 
  geom_smooth(formula = y ~ log(x), 
              method=glm, 
              se=FALSE, 
              color="red",
              size=0.5) +
  geom_point() +
  xlab("Habitat Suitability") +
  ylab("Observed Sex Ratio (F/M)") +
  theme_minimal() + xlim( c(0,1) ) + ylim(c(0.25,1.75))
```





