---
title: "Basic Graphing"
output: html_notebook
---

This notebook covers materials for basic plotting in R.

## Example Data Set

```{r}
summary(airquality)
```



```{r}
?airquality
```



## Univariate Plotting

Plotting of a single variable.

```{r}
temp <- airquality$Temp 
```



### Discrete Bins

Default plot for histograms.

```{r}
hist( temp )
```

Specifying where the breaks are for each bin.

```{r}
hist( temp, breaks = seq(50,100,by=5))
```

Plot as probability density instead of counting the frequency of each bin.

```{r}
hist( temp, freq = FALSE)
```


```{r}
hist( temp, axes=FALSE )
```



```{r}
p <- hist( temp, plot=FALSE )
p
```







### Density Estimation

Estimate a *density* of the data across the observed values as a continuous function.

```{r}
d <- density( airquality$Temp )
d
```

The object itself, is of 'class'

```{r}
class(d)
```

which knows how to plot itself.

```{r}
plot(d)
```


### Normality Plots

Testing to see if the data are 'normal'-ly distributed.

```{r}
qqnorm( temp )
qqline( temp, col="red")
```




## Bivariate Plotting



### Two Continuous Variables

Plotting of pairs of variables.

```{r}
plot(Wind ~ Solar.R, data=airquality)
```







### Continuous Response, Categorical Predictor

```{r}
summary(iris)
```



```{r}
boxplot( Petal.Length ~ Species, data=iris)
```





```{r}
mean_petal_length <- by( iris$Petal.Length, 
                         iris$Species, 
                         mean )
barplot(mean_petal_length,
        xlab="Species of Iris", 
        ylab="Average Petal Length")
```







## Kustomized Plotz

Scatter plot characters.


```{r}
plot( 1:25, 1:25, pch=1:25, cex=2, xlab="Plot Character", ylab="Plot Character" )
```


```{r}
plot( 1:25, 1:25, cex = (1:25)/10  )
```


Changing the axes & titles

```{r}
plot( Wind ~ Solar.R, data=airquality, bty="n", xlab="This is the x-axis label", ylab="The y-axis", main="Main Title" )
```


Scaling the range of both axes

```{r}
plot( Wind ~ Solar.R, data=airquality, xlim=c(0,500), ylim=c(0,50) )
```



Existing color palettes.  These values are in HEX (base 16 counting 0 - F) and have a pair of values for Red, a pair for Green, and a pair for Blue (e.g., #RRGGBB).  There are some built-in color palettes, each of which you specify the number of values.

```{r}
rainbow( n = 12 )
terrain.colors( n = 12 )
topo.colors( n = 12 )
```

We can plot the airquality data for these using the number of the month as the index on the rainbo

```{r}
plot( Wind ~ Solar.R, 
      pch = 19, 
      col=rainbow(12)[ Month ], 
      data=airquality)
```


Custom color spaces

```{r}
library(colorspace)
par(mfrow = c(3, 3))
cl <- sequential_hcl(5, "Heat")
for (i in c("map", "heatmap", "scatter", "spine", "bar", "pie", "perspective", "mosaic", "lines")) {
  demoplot(cl, type = i)
}
```

Colorbrewer (see also [colorbrewer website](http://colorbrewer2.org)).

```{r fig.height=10}
RColorBrewer::display.brewer.all()
```


Here is an example using *Set 3* from the Qualitative Set.

```{r}
boxplot( Solar.R ~ Month, data=airquality, col=brewer.pal(12, "Set3"))
```

### Barplots

```{r}
summary(mtcars)
```



```{r}
counts <- table( mtcars$gear )
counts
```

```{r}
barplot( counts,
         xlab="Number of Gears", 
         col = c("#FE1212", "#FED100", "#009B3A"))
```


## Pie Charts

Actually one of the worse types of plot for interpreting angles (humans suck at that task).

```{r}
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```


## Overlaying Objects

Cannot do this line-by-line, but have to do the whole chunk at one time.

```{r}
plot( Wind ~ Solar.R, data=airquality)

# absolute line with intercept = 15 and slope = 0, dashed and red.
abline(a=15, b=0, col="red", lty=2)

# abline with predicted line from regression as dotted blue
abline( lm( Wind ~ Solar.R, data=airquality), lty=3, col="blue")

# add a legend to the plot
legend( 75, 22, c("regression line", "fake line"), lty=c(3,2), col=c("blue","red"), bty="n", horiz=TRUE )
```

