---
title: "4 Projections"
author: "Your Name Here"
output: html_notebook
---


## Basic Background of Projections

Projections are a way that we can pretend to represent geospatial locations onto a 2-dimensional surface (paper or screen).  They are entirely constructs, made for us, because of the lower dimensional representation.

Here is a short video that discusses these issues *and* provides an example of how we can actually make maps that do not have distortions while at the same time can be viewed as a 2-dimensional represenation...  (intriguing I know...)

<iframe width="560" height="315" src="https://www.youtube.com/embed/D3tdW9l1690" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Projections in R

Perhaps one of the best ways for us to see this is to look at some maps and apply different projections to them.  I'm going to do this using the `maps` library^[If you do not have it, install it using `install.packages("maps")`].  

The projections being applied are detailed in:

```{r}
?mapproject
```

These are for demonstration purposes mostly, the key step in your data wrangling is that the first thing you should do is put all of your geospatial data into the same coordiante system and projections--and then forget about it!

### Equatorial Projections

These are projections centered on the prime meridian (e.g., Longitude=0).  This is a common projection we all are familiar with (and why Greenland looks bigger than Africa on some maps)

```{r}
library(maps)
map( "state", proj="mercator")
```

```{r}
# use gilbert

```

We can change the location of the prime meridian, here I'm moving the center of this projection to be in the center of the US.

```{r}
map("state", proj="cylequalarea", par=39.83)
```

But this sure distorts other parts of the globe, look at the size of Africa, S. America, and Australia.

```{r}
# change above to world instead of state

```

### Azimuth Projections

These projections are centered on the North Pole with parallels making concentric circles. Meridians are equally spaced radial lines.

Here are for

```{r}
par( mfrow = c(2,2),     # this bit of code makes 2x2 plots with no margin
     mar = c(0,0,0,0) ) 

# Plot 4 maps of the states with:
#  proj="azequidistant"
#  proj="fisheye", par=2
#  proj="perspective", par=8
#  proj="gnomonic"

```

### Polar Conic Projections

Here projections are symmetric around the Prime Meridian with parallel as segments of concentric circles with meridians being equally spaced.

```{r}
par( mfrow = c(2,2),     
     mar = c(0,0,0,0) )
map("state", proj="conic",   par = 39.83)
map("state", proj="lambert", par = c( 30, 40 ) )
map("state", proj="albers",   par = c( 30, 40 ) )
map("state", proj="simpleconic", par = c( 30, 40 ) )
```

### Silly Projections

There are some additional, miscellanouse projections, that we can play with.  

```{r}
par( mfrow = c( 2, 2 ), 
     mar = c( 0, 0, 0, 0) )

# do the above with:
#  proj="square"
#  proj="hex"
#  proj="bicentric", par=-98
#  proj="guyou"
```

## Coordinate Systems

Onto these projections, we can define individual coordinate systems.  These define the the physical location of entites (e.g., the coordinates).  There is a huge range of coordinate systems available for us to work with—sometimes I think people make up new ones just be snarky... That being said, the ones we will encounter often include:  
 - UTM (Universal Transverse Mercator) measuring the distance from the prime meridian for the x-coordinate and the distance from the equator (often called northing in the northern hemisphere) for the y-coordinate. These distances are in meters and the globe is divided into 60 zones, each of which is 6 degrees in width.   
 - Geographic coordinate systems use longitude and latitude. For historical purposes these are unfortunately reported in degrees, minutes, seconds, a temporal abstraction that is both annoying and a waste of time (IMHO).  
 - Decimal degrees, while less easy to remember, are easier to work with in R.  
 - State Planar coordinate systems are coordinate systems that each US State has defined for their own purposes. They are based upon some arbitrarily defined points of reference and another pain to use (IMHO). Given the differential in state area, some states are also divided into different zones. Maps you get from municipal agencies may be in this coordinate system. If your study straddles different zones or even state lines, you have some work ahead of you…  

It is best to use a system that is designed for your kind of work. Do not, for example, use a state plane system outside of that state as you have bias associated with the distance away from the origin. That said, Longitude/Latitude (decimal degrees) and UTM systems are probably the easiest to work with in R.

To play with these we will load in some example data, from Baja California.  Here are sampling locations in and the census of the number of males and female *Araptus attenuatus* individuals.  

This data file is a CSV file that is posted with this notebook.  Download it and put it into the `data` folder you have been working with thus far.

```{r}
araptus <- read.csv("data/araptus.csv")
```

These data look like the following and have a census of the number of male and female beetles collected at sites throughout Baja California as well as an estimation of the habitat suitability from a Niche Modeling exercise.

```{r}
summary( araptus )
```

The 'coordinates' we specified are just points, which we happened to label as "Latitude" and "Longitude" but we could have called them "X" and "Y" and unless you specify the CRS, R will pretend like it has no idea what system you are using.  If we want to make them represent real coordiantes (as in Latitude and Longitude we yanked off our GPS unit or grabbed from Google Earth), we need to specify the CRS directly.  

Here is a quick way to set a CRS using EPSG codes, an EPSG code (see [EPSG.io](https://epsg.io) for more information) is a shorthand notation for a coordinate reference system.  There are tons of different projects available and you can make up your own if you like.  For this example, I'll use one of the most common ones we will run into, `epsg:3857`.

```{r}
# apply to empty crs NOT A TRANSFORMATION

```

The `epgs:3857` is used by Google Maps, ESRI, OpenStreet Maps, Bing, and other providers and is equivallent to (for those who care) to the following `WKT` definition:

```
 PROJCS["WGS 84 / Pseudo-Mercator",
    GEOGCS["WGS 84",
        DATUM["WGS_1984",
            SPHEROID["WGS 84",6378137,298.257223563,
                AUTHORITY["EPSG","7030"]],
            AUTHORITY["EPSG","6326"]],
        PRIMEM["Greenwich",0,
            AUTHORITY["EPSG","8901"]],
        UNIT["degree",0.0174532925199433,
            AUTHORITY["EPSG","9122"]],
        AUTHORITY["EPSG","4326"]],
    PROJECTION["Mercator_1SP"],
    PARAMETER["central_meridian",0],
    PARAMETER["scale_factor",1],
    PARAMETER["false_easting",0],
    PARAMETER["false_northing",0],
    UNIT["metre",1,
        AUTHORITY["EPSG","9001"]],
    AXIS["X",EAST],
    AXIS["Y",NORTH],
    EXTENSION["PROJ4","+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext  +no_defs"],
    AUTHORITY["EPSG","3857"]]
```

with the following `proj4` string

```{r}
library( rgdal )

make_EPSG()  %>%
  dplyr::filter( code == 3857) %>% 
  dplyr::select( prj4 ) 
```


### The Shortcut

OK, now that we know a little about projections and the `sf` library, we can skip all of these intermediate steps and collapse them into a single call.  Remember, object we are using is a `data.frame` to begin with.

```{r}
class( araptus )
```

Going forward, we can load in a data set, make it into an `sf` object, and apply specify the coordinate reference system quickly as:

```{r}
# read in and translate as one.
```

It is just that we had to go through all the intermediate steps so you could understand what is going on in this single import step.

## Transforming Points

So perhaps we want to transform the coordinates from Longitude/Latitude to one of the coverages specifically designed for Mexico.  To do this, we need to find a definition of the new projection we are using.  For simplicity, I will translate the Lat/Lon of the previous data set into UTM using Zone 11 which covers the west coast of north america.  This happens to be espg code 32612 and you can find it by searching [EPSG.io](EPSG.io).

```{r}
make_EPSG()  %>%
  dplyr::filter( code == 32611) %>% 
  dplyr::select( prj4 ) -> mex_proj

mex_proj
```

So we can use either the epsg code:

```{r}
# transform to EPSG:32611

```

or the full proj4 string as:

```{r}
st_transform(arapat_sf,crs = mex_proj$prj4[1] ) -> arapat_mex
arapat_mex
```


Now, when we plot it, check out the axes.

```{r}
# plot the data with axes
```

### The `sf` Library

There is a very useful library for us to play with using what is called *Simple Features*.  This library defines some new data types for POINTS, LINES, POLYGONS, etc that can be used in normal `data.frame` objects (rather than hiding the data in odd structures like the `sp` library does).

Here I take the above data and indicate that a column will be treated as a geometry object representing both Longitude and Latitude.  This makes it *both* a `data.frame` and a `sf` object simultaneously.

```{r}
library( sf )
library(magrittr)  # must have this to prevent namespace collisions


```

And if we look at it, it looks like:

```{r}
# Look, a geometry object as a SINGLE column in a data.frame!
```

and individual rows are:

```{r}
# print it out and look
```

Notice how the geometry column is recognized as a POINT object.  This is *critical* as it is the *lingua Franca* for interacting with GeoDatabase objects, online services, etc.

One of the things we need to shore up is that the points here have no coordinate reference system.  

```{r}
# show CRS
```

### GGPlot and `sf` Objects

Just because we have now simplified our data representation to have `geometry` objects represented within a `data.frame` with the rest of the data, we should now be able to leverage the power of ggplot2 to make our lives even better...  

So when we pass ggplot a `data.frame` that has `geometry` objects as columns, it will automagically use those for coordinates.  As a result, we only need to specify data for the geometric layer we want to overlay on it and how to represent those data on the plot.  

In this example, we have a set of points in the `geometry` that we are able to use to show differences in the sex ratio (the ratio of females to males) of the locale and then use this to plot it.

```{r}
library(ggplot2)  # Our old friend

```


## In Class Activities

Take the `arapat` data set and tranform it into another projection. 

```{r}
# arapat -> some other projection found on epsg.io

```


Create a plot where habitat suitability is displayed spatially for the transformed data set.

```{r}
# Habitat suitability 

```















