---
title: "7 dplyr Basics"
output: html_notebook
---

For this, we need to get a relatively large data file so we can start playing with it.  I'm going to use some data that comes with the `nycflights13` library.  If you do not have it, you will need to install it.

```{r eval=FALSE}
install.packages("nycflights13")
```

and then load it in and we can look at the `flights` data.frame.

```{r}
library(nycflights13)
data(flights)  # this loads the data from the package
summary( flights )
```

These data consist of `r formatC(nrow(flights), big.mark=",")` records for `r ncol(flights)` variables (named *`r paste(names(flights), collapse=", ")`*).  That is a *reasonable* sized dataset to work with.  

See `?flights` for a description of the data contained within.

## Overview of `dplyr`

The `dplyr` library is one of the libraries included in the `tidyverse` constellation of libraries.  The main focus for `dplyr` is to allow you to manipulate data frames in the following ways:  

- Pick certain rows of data based upon their values (use `filter()`)  
- Choose particular subsets of columns or rearrange them (use `select()`)  
- Create new columns of data (use `mutate()`)
- Sort data on one or more columns (use `arrange()`)
- Group data based upon specific columns (use `group_by()`)  
- Summarize columns of data or composites (use `summarize()`)


Using the piping we've been introduced to thus far via `%>%` we can chain these kinds of operations together into single lines of code that perform rather complicted patterns of communication and analysis.  

Let's load the library in and start playing around a bit.

```{r message=FALSE, warning=FALSE}
library(dplyr)
```


## Filter

The `filter()` function allows us to select particular rows from the data set.  To specify the criteria of the filter, we use *logical* operations (things like `>`, `>=`,  `==`, `<=`, `<`, `is.na()`, `is.null()`, etc.).  

Here are all the flights that left a NYC airport and went to Anchorage Alaska. 

```{r}
flights %>%
  filter( dest == "ANC" )
```

Here is a look at the number of flights whose flight time (measured as minutes in the `air_time` variable) was more than 10 hours in length.

```{r}
flights %>%
  filter( air_time > (60 * 10) ) %>%
  nrow() 
```

We can do compound expressions for `filter()` (and the rest) by adding addition criteria.  Here I sort by both month (January) and carrier (AS for *Alaska Airlines*) as either calling `filter()` twice (e.g., the output of one become the input of another)

```{r}
flights %>%
  filter( month == 1 ) %>%
  filter( carrier == "AS" )
```

or as a single call to filter with the statements separated by a comma (this time I search for *Mesa Airlines* in May).  Both approaches give the same results as they are conducting an **and** operation (e.g., the conditions for both *must* be `TRUE`)

```{r}
flights %>%
  filter( month == 5, 
          carrier == "YV" )
```

### Complex Filters

We can combine additional logical operators in the filter to make more complex filters.  The code below does the following filter (if reading it out loud it would be)
1. from all the flights, 
2. find ones that are in January or Feburary, 
3. whose destination is Seattle or Portland (I use the OR operator `|` for both destinations here), 
4. and airlines that were neither *Alaska* nor *United* (I use the negation `!` and set  `%in%` operators for that one).

```{r}
flights %>%
  filter( month < 2, 
          dest == "SEA" | dest == "PDX",
          !(carrier %in% c("AS","UA")) )
```



## Selecting Subsets of Columns

In additiont to filtering rows, we can select subsets of columns.  We do this, surprisingly, with the function `select()`.  To select individual columns, give their variable names (not quoted).

```{r}
flights %>%
  select( year, month, day, dest, air_time ) 
```


To select *NOT* a column, use the name preceeded by a negative sign (e.g., a dash) and it will omit that column.

```{r}
flights %>%
  select( -year, -month, -day ) 
```

You can also use the `select()` to re-order the columns, and this is especially helpful if you pair this with the `everything()` function that give (drum roll please), everything else (way helpful if you have a lot of columns of data and do not want to type them all out).

```{r}
flights %>%
  select( dep_time, air_time, carrier, everything() )
```

You can also use the select to rename the columns as you grab them.  This is really nice for setting up your variable names when you do something like making a new plot.  Grab the data, pipe it through the select and give it names that will be approprite for axis labels and then pass it on to `ggplot` and you won't have to do the `+ xlab() + ylab()` fixes for odd variable names.

```{r}
flights %>%
  select( Airline = carrier, 
          Departure = sched_dep_time,
          Arrival = sched_arr_time,
          Duration = air_time )
```

If you want to keep all the data and just rename a couple of columns, use the `rename()` function and it will change them in place.

```{r}
flights %>%
  rename(Departure = dep_time)
```





## Creating New Variables with `mutate()`

In the process of piping the data, we can add to the `data.frame` by using `mutate()`.  Here is an example where I take the *month*, *day*, and *year* columns and paste them togehter using the slash (`/`) character and then turn it into a *real* date object.  I then select the date, destination, flight time (notice how I use the back-tick quote to make a name that has a space in it), and Delay.


```{r}
library( lubridate )
flights %>%
  mutate( Date = parse_date_time( paste( month, day, year, sep="/"),  
                                  orders="%m/%d/%Y") ) %>%
  select( Date, 
          Destination = dest,
          `Flight Time` = air_time, 
          Delay = arr_delay) 
```

Just like filter, I can chain these together to do multiple mutations^[Careful, if you do a mutation that makes a new column, you cannot refer to this new column in the same mutation, you will have to make a subsequent mutation to reference the newly created column].  Here I do the same as above but also make the destination a categorical `factor` variable.

```{r}
flights %>%
  mutate( Date = parse_date_time( paste( month, day, year, sep="/"),  
                                  orders="%m/%d/%Y") ) %>%
  mutate( Destination = factor( dest ) ) %>%
  select( Date, 
          Destination,
          `Flight Time` = air_time, 
          Delay = arr_delay) %>%
  summary()
```


## Sorting 

We can arrange the data using `arrange()` in either increasing or decreasing orders.  Here are the shortest flights from NYC.

```{r}
flights %>%
  select( Destination = dest, 
          Duration = air_time ) %>%
  arrange( Duration )
```

and the longest flights (converted to length in hours).  To switch from largest to smallest, you can use `desc(Duration)` or just put a negative symbol in front of the variable name.

```{r}
flights %>%
  mutate( Duration = air_time / 60.0 ) %>%
  select( Destination = dest, 
          Duration ) %>%
  arrange( -Duration ) 
```


## Summaries 

So all of the stuff thus far has been selecting rows and columns of data and changing entire columns, now we start talking about summarizing the data we have manipulated.  

```{r}
flights %>%
  mutate( `Total Delay` = dep_delay + arr_delay ) %>%
  summarize( `Departure Delay` = mean(dep_delay),
             `Arrival Delay` = mean( arr_delay),
             `Total Delay` = mean( `Total Delay`) )
```

OK, notice one thing, it is all `NA` values! That is because there are missing data and *by default* `R` will return a `NA` for an operation if any of the data are `NA`.  This is a "good thing"^&copy;^ as it alerts you, the researcher, that there is missing data and you will have to deal with it to get a usable answer.  There are two ways to try this, we could filter out all the missing data first using `filter()`


```{r}
flights %>%
  mutate( `Total Delay` = dep_delay + arr_delay ) %>%
  filter( !is.na(`Total Delay`) ) %>%
  summarize( `Departure Delay` = mean(dep_delay),
             `Arrival Delay` = mean( arr_delay),
             `Total Delay` = mean( `Total Delay`) )
```

**OR** we could tell the `mean()` function to ignore missing stuff.

```{r}
flights %>%
  mutate( `Total Delay` = dep_delay + arr_delay ) %>%
  summarize( `Departure Delay` = mean(dep_delay, na.rm = TRUE),
             `Arrival Delay` = mean( arr_delay, na.rm = TRUE),
             `Total Delay` = mean( `Total Delay`, na.rm = TRUE ) )
```


We can also `group_by()` to get more rows in our output summary.  Following from the previous example, if we want the data on worst arrival delays by destination, we can

```{r}
flights %>%
  group_by( Destination = dest ) %>%
  summarize( Delay = mean( arr_delay, na.rm=TRUE ),
             Distance = mean( distance, na.rm=TRUE),
             `Late Flights` = length( arr_delay > 0 ) ) %>%
  arrange( -Delay ) 
```

Which shows that the following are the worst Columbia Metropolitan Airport in South Carolina had the longest average delay of all destinations, averaging about 41 minutes.  

The object that pops out the end of the `summarize()` function is another `data.frame` object, so we can actually just pipe this into a `ggplot` object.  Here are the destination cities whose late arrival averages over the entire year exceeds 20 minutes.

```{r message=FALSE, warning=FALSE}
library( ggplot2 ) 
library( ggrepel )


flights %>%
  group_by( Destination = dest ) %>%
  summarize( Delay = mean( arr_delay, na.rm=TRUE ),
             Distance = mean( distance, na.rm=TRUE),
             `Daily Late Flights` = length( arr_delay > 0 )/365 ) %>%
  mutate( Label = ifelse( Delay > 20, Destination, NA ) ) %>%
  ggplot( aes(Distance, Delay) )  + 
  stat_smooth( se = FALSE, 
               col="red", 
               linetype = 2) +
  geom_point( aes(size = `Daily Late Flights`), 
              alpha=0.75) + 
  geom_label_repel( aes(label=Label) ) + 
  theme_bw() 

```

Combining these with the `knitr` and `knitrExtras` packages, we can easily make tables in from our data that can be inserted into the document.  Here is a version that you can 

```{r}
library(knitr)
library(kableExtra)

flights %>%
  group_by( Destination = dest ) %>%
  summarize( Delay = mean( arr_delay, na.rm=TRUE ),
             Distance = mean( distance, na.rm=TRUE),
             `Daily Late Flights` = length( arr_delay > 0 )/365 ) %>%
  arrange( -Delay ) %>%
  head() %>%
  kable( 'html' ) %>%
  kable_styling( bootstrap_options = "striped", 
                 full_width = FALSE ) 
```




















