---
title: "Basic Graphing Homework"
author: "YOUR NAME HERE"
output: html_notebook
---

## The Data

For this assignment, we will be using the built-in data from the annual review for 1974 from *Motor Trends Magazine*.  Here are the columns for the raw data in that dataset (it is `mtcars` in R).

```{r}
names(mtcars)
```

These don't mean much (see `?mtcars` for a description) and we don't need them all for this homework.  I'll take a subset of the data to play with and clean it up in the code below (it is OK if it looks a bit puzzling, in a week or so, you'll totally be able to do this).  I call the new data `data` (I know, very imaganitive).

```{r message=FALSE }
library(tidyverse)

mtcars %>%
  select( MPG = mpg, 
          Displacement = disp, 
          Horsepower = hp, 
          Weight = wt, 
          QuarterMile = qsec, 
          Cylinders = cyl, 
          Gears = gear, 
          Transmission = am ) %>%
  mutate( Gears = factor( Gears, ordered=TRUE ), 
          Transmission = factor( ifelse( Transmission == 0, "Automatic", "Manual") ), 
          Cylinders = factor( Cylinders, ordered=TRUE ) ) -> data

summary(data)
```

**IMPORTANT:** In all the following plots, make sure you label the axes appropriately and make the plot *look like* it is supposed to be in a publication.  


## Univariate Plots

1. Make a plot of the distrubtion for each of the variables MPG.  

```{r}
hist( data$MPG, xlab = "Miles Per Gallon", main="") 
```

2. Make a Density plot of the Displacement.

```{r}
d <- density( data$Displacement )
plot(d, bty="n", main="", xlab="Engine Displacement")
```


3. Create a scatter plot of MPG as a function of Horsepower.  Color each symbol by the number of cylinders and make the shape of the point indicate the transmission type.

```{r}
plot( MPG ~ Horsepower, 
      data=data, 
      ylab="Miles Per Gallon (MPG)",
      col = Cylinders, 
      pch = as.numeric(Transmission) , 
      bty = "n")

legend( 150, 30, c("Automatic","Manual"), 
        pch=c(1,2), 
        bty="n", 
        horiz = TRUE)

legend( 150, 28, c("3 Cylinder","4 Cylinder","5 Cylinder"), 
        col=c("black","red","green"), 
        pch=20, 
        horiz = TRUE, 
        bty="n")
```


