---
title: "Homework for `dplyr`"
author: "YOUR NAME HERE"
output: html_notebook
---

This is a short set of homework for using the `dplyr` library.  It is going to use the `flight` data, we played with in the lecture content.

```{r}
library(tidyverse)
library(nycflights13)
data(flights)
```


1. What is the average delay for each month during 2013 of flights from?  Make the ouptut into a nicely formated table and use month names.

```{r}
library( lubridate)

flights %>%
  filter( dest == "RIC" ) %>%
  mutate( Delay = arr_delay + dep_delay ) %>%
  mutate( Date = parse_date_time( paste(month,day,year,sep="/"), 
                                  orders="%m/%d/%Y") ) %>%
  mutate( Month = months(Date)) %>%
  select( Month, Delay) %>%
  group_by( Month ) %>%
  summarize( `Flight Delay` = mean( Delay , na.rm = TRUE) ) %>% 
  knitr::kable('html') %>%
  kableExtra::kable_styling( bootstrap_options = c( "striped", 
                                                    "hover",
                                                    "condensed"),
                             full_width=FALSE)
```

2. Which day of the week are you more likely to *arrive on time* to Phoenix Arizona?  Make the output into a bar plot.

```{r}
flights %>%
  filter( dest == "PHX" ) %>%
  mutate( Arrival = arr_delay ) %>%
  mutate( Date = parse_date_time( paste(month,day,year,sep="/"), 
                                  orders="%m/%d/%Y") ) %>%
  mutate( Weekday = factor( weekdays(Date), 
                            ordered=TRUE,
                            levels = c("Monday",
                                      "Tuesday",
                                      "Wednesday",
                                      "Thursday",
                                      "Friday",
                                      "Saturday",
                                      "Sunday") ) ) %>%
  select( Weekday, Arrival ) %>%
  group_by( Weekday ) %>%
  summarize( Arrival = mean(Arrival, na.rm=TRUE) ) %>%
  ggplot( aes(x=Weekday, y= Arrival) ) + 
  geom_col() + ylab("Average Arrival Delay") + 
  theme_bw()
```

3. Which airlines have, on average, arrival times that are earlier than scheduled?


```{r}
flights %>%
  select( carrier, arr_delay ) %>%
  group_by( `Airline` = carrier ) %>%
  summarize( `Average Delay` = mean( arr_delay, na.rm=TRUE ) ) %>%
  arrange( `Average Delay` ) %>%
  filter( `Average Delay` <=  0 ) %>%
  knitr::kable('html') %>%
  kableExtra::kable_styling( bootstrap_options = c( "striped", 
                                                    "hover",
                                                    "condensed"),
                             full_width=FALSE)
```

4. What are the five airports that have the shortest flight times?  Give the results as a table.


```{r}
flights %>%
  arrange( air_time ) %>%
  group_by( dest ) %>%
  summarize( `Flight Time` = min( air_time, na.rm=TRUE ) ) %>%
  select( `Flight Time`, 
          `Destination` = dest) %>%
  arrange( `Flight Time` ) %>%
  head( n = 5 ) %>%
  knitr::kable('html') %>%
  kableExtra::kable_styling( bootstrap_options = c( "striped", 
                                                    "hover",
                                                    "condensed"),
                             full_width=FALSE)
```

5. Is there a correlation between flight distance and total delay?  Use the function `cor.test( X, Y )`.  Interpret the output.

```{r}
flights %>%
  mutate( Delay = arr_delay + dep_delay ) %>%
  select( Distance = distance, 
          Delay ) -> tmp

cor.test( tmp$Distance, tmp$Delay )
```

```{r message=FALSE, warning=FALSE }
ggplot( tmp, aes(Distance, Delay) ) + 
  geom_point(  )  + 
  stat_smooth( se=FALSE )
```



