---
title: "Getting Data Into R"
author: 'Import Workflows'
date: 'ENVS543: Environmental Data Literacy'
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    hightlight: tango
    logo: "../../logo.png"
    css: "../../style.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warking=FALSE, error = TRUE, message = FALSE)
options("kableExtra.html.bsTable" = T)
```

## Impetus

One of the first things we do each time we use R is to import or load data into the `Environment`.  These data may be:

 - Input directly in R  
 - Loaded from the local filesystem  
 - Available online via a URL/URI  
 - Within a database  
 


# Direct Input of Data

## Direct Input - From Vectors

You can easily input data into R within your scripts or on the command line.

```{r}
x <- c(1,24,4,65,3,2,14)
y <- rnorm(7)
df <- data.frame( X = x, 
                  Y = y )
df
```


## Direct Input - At Once

```{r}
df <- data.frame( X = c(1,24,4,65,3,2,14),
                  Y = rnorm(7) )
df
```


## Tidyverse

The `tidyverse` library has a ton of extentions to `base R` that make data wrangling much more fluid and efficient.  The sooner you start using this constellation of packages, the quicker you will begin emjoying your life... 

```{r}
library(tidyverse)
```


## Parsing Vectors

The vector is the base container for all data types in `R`.  The `readr` library (within `tidyverse`) has several `parse_*` functions that can assist in importing data types.

### Converting String to Numeric
```{r}
parse_integer( c("2","5",NA,"52"))
parse_integer( c("2","5",".","52"), na = ".")
```


## Parsing Errors {.build}

If the `parse_*` function runs into a problem coercing an entry into a specified data type, it will tell you.

```{r}
dr_d <- parse_integer( c("23","42","Bob"))
```

And give you some additional feedback if you ask it what problems it had.

```{r}
problems( dr_d )
```



## Tibbles

A `tibble` is just like a `data.frame` object and can often be used in the same way.  You can use the `read_csv` function to load in a single string that has all the data contained within it.  The `read_csv` will attempt to guess the best kind of data to input into the structure.

```{r}
df <- read_csv( "Name, Value, Passed
A, 63.4, TRUE
B, 54.2, FALSE
C, 92.4, TRUE")
df
```




# Data On Your Computer




# Data Online



# Data Within Databases







