---
title: "Models"
author: 'Models Workflows'
date: 'ENVS543: Environmental Data Literacy'
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    hightlight: tango
    logo: "../../logo.png"
    css: "../../style.css"
---

```{r setup, include=FALSE}
source("../../../options.R")
```

## Models {.build}

> Now it would be very remarkable if any system existing in the real world could be exactly represented by any simple model. However, cunningly chosen parsimonious models often do provide remarkably useful approximations. For example, the law $PV = RT$ relating pressure $P$, volume $V$ and temperature $T$ of an “ideal” gas via a constant $R$ is not exactly true for any real gas, but it frequently provides a useful approximation and furthermore its structure is informative since it springs from a physical view of the behavior of gas molecules.
>
> For such a model there is no need to ask the question  `Is the model true?` If “truth” is to be the “whole truth” the answer must be “No.” The only question of interest is  `Is the model illuminating and useful?`
>
>The goal of a model is not to uncover truth, but to discover a simple approximation that is still useful.”


Often shortened to the quote  `All models are wrong, but some are useful.`




## Example Data

```{r echo=FALSE}
load("../../data/araptus.rda")
```

```{r}
araptus %>% 
  mutate( SexRatio = Females/Males ) -> araptus 
summary( araptus )
```

## Habitat Suitability

```{r}
araptus %>%
  ggplot( aes(Suitability, SexRatio)) + geom_point() + 
  stat_smooth(method="lm", se=FALSE, col="red", size=0.5) + 
  xlim(c(0,1)) + xlab("Habitat Suitability") + ylab("Observed Sex Ratio")  
```

## Fitting a Model

We have an infinite number of models that could be fit to these data.  Here are some random examples.
```{r}
models <- data.frame( intercept = runif(500, -3, 3),
                      slope = runif(500, -3, 3 ) )
head( models )
```



## Potential Models

```{r}
ggplot(araptus, aes(Suitability, SexRatio)) + 
  geom_abline( aes(intercept=intercept, slope=slope), data=models, alpha=0.25 ) +
  geom_point() + xlim(c(0,1)) + xlab("Habitat Suitability") + ylab("Observed Sex Ratio")
```


## Least Square Distances

One way to figure out which of these models is `preferred` is to find the line which minimizes the distance from all the points to the predicted line.  
```{r echo=FALSE}
araptus %>%
  mutate( SexRatio1 = araptus$Suitability*0.4 + 0.75) %>%
  ggplot( aes(Suitability,SexRatio) ) + 
  geom_point() + xlim(c(0,1)) + 
  geom_abline( intercept = 0.75, slope = 0.4, alpha=0.25 ) +
  geom_linerange( aes(ymin=SexRatio,ymax=SexRatio1), color="blue" ) +
  xlab("Habitat Suitability") + ylab("Observed Sex Ratio")
```

## Function To Find Model Least Squares

Let's take a circuituous route there and work through some functions.   
```{r}
pred <- function( df, intercept, slope ) {
  return( intercept + slope * df[,1] )
}

araptus %>%
  select( Suitability, SexRatio ) -> df

Yhat <- pred(df, models$intercept[1], models$slope[1] )
Yhat

```


## Function To Find Model Least Squares {.build}

```{r}
squareDist <- function( coefs, X, Y ) {
  Yhat <- coefs[1] + coefs[2]* X
  d <- (Y - Yhat)^2
  return( sum( d ) )
}
```

```{r}
models$SumOfDistances <- apply( models, 
                      1, 
                      function(coefs, df) {
                        return( squareDist(coefs, df$Suitability, df$SexRatio) )
                        }, 
                      df=df)
```

## Evaluations of Models

```{r}
models %>%
  ggplot( aes(intercept, slope, alpha = 1.0/SumOfDistances)) + 
  geom_point() + xlab("Model Intercept") + ylab("Model Slope") + 
  theme( legend.position = "none")
```

## A Mesh View

```{r echo=FALSE}
library(plotly)

plot_ly( x = ~ models$slope, 
         y = ~ models$intercept, 
         z = ~ models$SumOfDistances, type="mesh3d",
         opacity = 0.8
         )
```


